\documentclass[12pt]{amsart}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue
}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{question}[theorem]{Question}
\newtheorem{caution}[theorem]{Caution}

\begin{document}

\title{Homework 1}
\author{Group 2}

\maketitle

\begin{enumerate}
	\item (1.2) Decrypt each of the following Caesar encryptions by trying the various 
	possible shifts until you obtain readable text.
	\begin{itemize}
		\item \texttt{LWKLQNWKDWLVKDOOQHYHUVHHDELOOERDUGORYHOBDVDWUHH}
		\item \texttt{UXENRBWXCUXENFQRLQJUCNABFQNWRCJUCNAJCRXWORWMB}
		\item \texttt{BGUTBMBGZTFHNLXMKTIPBMAVAXXLXTEPTRLEXTOXKHHFYHKMAXFHNLX}
	\end{itemize}
	
	\medskip

	\textit{Solution.}

	I deciphered all of the following by hand on an iPad.
		\begin{itemize}
			\item \texttt{LWKLQNWKDWLVKDOOQHYHUVHHDELOOERDUGORYHOBDVDWUHH} \\
                This cipher took 3 shifts. \\
                \texttt{I THINK THAT I SHALL NEVER SEE A BILLBOARD LOVELY AS A TREE}
			\item \texttt{UXENRBWXCUXENFQRLQJUCNABFQNWRCJUCNAJCRXWORWMB} \\
                This cipher took 9 shifts. \\
                \texttt{LOVE IS NOT LOVE WHICH ALTERS WHEN IT ALTERATION FINDS}
			\item \texttt{BGUTBMBGZTFHNLXMKTIPBMAVAXXLXTEPTRLEXTOXKHHFYHKMAXFHNLX}\\
                This cipher took 19 shifts. \\
                \texttt{IN BAITING A MOUSE TRAP WITH CHEESE ALWAYS LEAVE ROOM FOR THE MOUSE}
		\end{itemize}

  \bigskip

  \item (1.3) Use the simple substitution table below
  \begin{center}
	  \begin{tabular}{|c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c| c| c|}
		  \hline
		  a & b & c & d & e & f & g & h & i & j & k & l & m & n & o & 
		  p & q & r & s & t & u & v & w & x & y & z \\
		  \hline
		  S & C & J & A & X & U & F & B & Q & K & T & P & R & W & E & 
		  Z & H & V & L & I & G & Y & D & N & M & O \\
		  \hline
	  \end{tabular}
  \end{center}
  \begin{enumerate}
	  \item Encrypt the plaintext message
	  \begin{center}
		  \texttt{The gold is hidden in the garden.}
	  \end{center}
	  \item Make a decryption table, that is, make a table in which the ciphertext 
		  alphabet is in order from A to Z and the plaintext alphabet is mixed up.
	  \item Use your decryption table from (b) to decrypt the following message.
	  \begin{center}
		  \texttt{IBXLX JVXIZ SLLDE VAQLL DEVAU QLB}
	  \end{center}
  \end{enumerate}
  
  \medskip
  \textit{Solution.}

 	 \begin{enumerate}
		\item The encrypted message is
		\begin{center}
			\texttt{IBXFE PAQLB QAAXW QWIBX FSVAX W}
		\end{center}
	 	\item The decryption table for the given cipher:
            \begin{center}
                \hspace*{-1cm}
    			\begin{tabular}{|c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c| c| c|}
    				\hline
    				A & B & C & D & E & F & G & H & I & J & K & L & M & N & O & 
    				P & Q & R & S & T & U & V & W & X & Y & Z \\
    				\hline
    				d & h & b & w & o & g & u & q & t & c & j & s & y & x & z & 
    				l & i & m & a & k & f & r & n & e & v & p \\
    				\hline 
    			\end{tabular} 
	        \end{center} 
            \medskip
            \item \texttt{THE SECRET PASSWORD IS SWORDFISH} 
	 	\end{enumerate}

\bigskip

\item (1.4.c) Each of the following messages has been encrypted using a simple
	substitution cipher. Decrypt them. For your convenience, we have given
	you a frequency table and a list of the most common bigrams that appear
	in the ciphertext. (If you do not want to recopy the ciphertexts by
	hand, they can be downloaded or printed from the web site listed in the
	preface.) In order to make this one a bit more challenging, we have
	removed all occurrences of the word “the” from the plaintext. 

	``A Brilliant Detective"
	\begin{center}
		\ttfamily
		GSZES GNUBE SZGUG SNKGX CSUUE QNZOQ EOVJN VXKNG XGAHS AWSZZ
		BOVUE SIXCQ NQESX NGEUG AHZQA QHNSP CIPQA OIDLV JXGAK CGJCG
		SASUB FVQAV CIAWN VWOVP SNSXV JGPCV NODIX GJQAE VOOXC SXXCG
		OGOVA XGNVU BAVKX QZVQD LVJXQ EXCQO VKCQG AMVAX VWXCG OOBOX
		VZCSO SPPSN VAXUB DVVAX QJQAJ VSUXC SXXCV OVJCS NSJXV NOJQA
		MVBSZ VOOSH VSAWX QHGMV GWVSX CSXXC VBSNV ZVNVN SAWQZ ORVXJ
		CVOQE JCGUW NVA
	\end{center}
		
	The ciphertext contains $313$ letters. Here is a frequency table: 
	\begin{center}
		\ttfamily
		\begin{tabular}{|c||c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c| c| c|}
			\hline
			& V & S & X & G & A & O & Q & C & N & J & U & Z & E & W
			& B & P & I & H & K & D & M & L & R & F \\
			\hline
			Freq  & 39 & 29 & 29 & 22 & 21 & 21 & 20 & 20 & 19 & 13
			      & 11 & 11 & 10 & 8 & 8 & 6 & 5 & 5 & 5 & 4 & 3 &
			2 & 1 & 1 \\
			\hline
		\end{tabular}
	\end{center}

	The most frequent bigrams are: \texttt{XC} (10 times), \texttt{NV} (7
	times), and \texttt{CS}, \texttt{OV}, \texttt{QA}, and \texttt{SX} (6
	times each).

	\medskip
	\textit{Solution.}
	
	After a fair amount of trial and error (including confusion as to why the string ``the'' was appearing and then realizing it was a substring of a word) and looking at common bigrams and trigrams, we noticed that ``XCSXXC'' appears 3 times. Comparing this with the common bigrams, we tried assigning this string to ``thatth'', and then assigning `V' to `e' since it was the most frequent and followed ``thatth'' twice. We then used the letter frequencies to start tentatively adding in other vowels, and by this point enough was filled in that it was fairly clear how to fill in the rest of the letters based on strings looking like certain words. We believe this is the decryption key (excluding `T' and `Y' since they never appear):

	\begin{center}
		\begin{tabular}{|c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c|}
			\hline
			A & B & C & D & E & F & G & H & I & J & K & L & M & N & O & 
			P & Q & R & S & U & V & W & X & Z \\
			\hline
			n & y & h & b & f & z & i & g & u & c & w & j & v & r & s & 
			p & o & k & a & l & e & d & t & m \\
			\hline 
		\end{tabular}
	\end{center}

	Thus, the decrypted message (with adding instances of ``the'' where they seem to have been removed) is:

	\begin{center}
		\ttfamily
		I am fairly familiar with all forms of secret writing, and am myself the author of a trifling monograph upon the subject, in which I analyze one hundred separate ciphers, but I confess that this is entirely new to me. The object of those who invented this system has apparently been to conceal that these characters convey a message, and to give the idea that they are mere sketches of children.
	\end{center}


	\bigskip
	\item (1.5) Suppose that you have an alphabet of 26 letters.
		\begin{enumerate}
			\item  How many possible simple substitution ciphers are there?
			\begin{center}
				\texttt{26! Different simple substitution. 26 possibilities for the first and decreases as a letter is taken.}
			\end{center}
			\item  A letter in the alphabet is said to be fixed if the encryption of the letter is the
				letter itself. How many simple substitution ciphers are there that leave:
			\begin{enumerate}
				\item No letters fixed?
				\begin{center}
					\texttt{26! Since no letters can't be themselves}
				\end{center}
				\item At least one letter fixed?
				\begin{center}
					\texttt{26! - 25! Total different minus No fixed letters with at least one}
				\end{center}
				\item Exactly one letter fixed? 
				\begin{center}
					\texttt{25! With only one letter fixed on itself. }
				\end{center}
				\item At least two letters fixed?
				\begin{center}
					\texttt{26! - 25! -24! + 23! }
				\end{center}
			\end{enumerate}
			(Part (b) is quite challenging! You might try doing the problem first with an alphabet 
			of four or five letters to get an idea of what is going on.)
		\end{enumerate}

	\end{enumerate}

\end{document}
